---
format:
  pdf:
    tbl-cap-location: top
    dpi: 500
    fig-pos: "H"
    number-sections: true
    cite-method: natbib
    include-in-header: 
      - text: |
          \usepackage{lscape}
          \newcommand{\blandscape}{\begin{landscape}}
          \newcommand{\elandscape}{\end{landscape}}
          \usepackage{rotating}
      - file: preamble.tex 
    include-before-body: title.tex
    keep-tex: TRUE
  docx:
    reference-doc: new_template_no_numbering.docx
pdf-engine: xelatex
# monofont: "DejaVu Sans Mono"
# mainfont: "Times New Roman"
execute: 
  echo: false
  warning: false
editor: source
bibliography: references.bib
biblio-style: apalike
natbiboptions: round
link-citations: true
linkcolor: blue!000!black
urlcolor: blue!000!black
citecolor: blue!000!black
linestretch: 1.5
indent: TRUE
title-block-banner: false
title-block-style: none
---

```{r setup}
#| include=FALSE
options(scipen=999)
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      dpi = 500, 
                      fig.pos = "H", 
                      fig.env= 'figure', 
                      tab.topcaption = FALSE,
                      ft.align = "left",
                      ft.arraystretch = 0.5                      
                      )
library(tidyverse)
library(readr)
library(readxl)
library(zoo)
library(magick)
library(stringi)
library(patchwork)
library(xtable)
library(scales)
library(rvest)
library(rebus)
library(PNWColors)
library(lubridate)
library(here)
library(flextable)
library(gt)
library(officer)
library(officedown)
library(modelsummary)
library(sandwich)
library(lmtest)

set_flextable_defaults(
  font.size = 8,
  digits = 2,
  font.family = "Arial",
  border.color = "#5d74a5"
)


```

```{r functions}
source(here("Functions","flextable_word.R" ))
source(here("Functions","modelsummary_word.R" ))
source(here("Functions","modelplot_word.R" ))
source(here("Functions","borderlines.R" ))
```


# Introduction


# Literature Review
@alexakis2023animal

# Data and Methodology
\newpage

```{r}
#| label: tbl-descriptive
#| tbl-cap: "Descriptive Statistics"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 0.5

descriptives <- read_rds(here("Outputs", "combined_data", "artifacts_descriptives.rds"))
descriptives_tbl <- descriptives$descriptives_tbl

descriptives_tbl %>%
  as_grouped_data(groups = c("Group")) %>% 
  as_flextable(hide_grouplabel = TRUE) %>% 
  colformat_double(digits = 2, i = 2:21, j = 2:6) %>% 
  colformat_double(digits = 0, i = 22:46, j = 2:6) %>% 
  colformat_double(digits = 2, i = 22:46, j = 3) %>% 
  width(j = 1, width = 1.5) %>%
  align(align = "center", i = c(1, 11, 21, 24), part = "body") %>% 
  autofit(add_w = 0)
```

# Results

## Macroprudential regulation
\blandscape

```{r}
#| label: tbl-macropru_draft_lending
#| tbl-cap: "Macroprudential regulations and lending volumes (3-months) results"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

macropru_results <- read_rds(here("Outputs", "models", "artifacts_macropru_models.rds"))

macropru_three_month_results <- read_rds(here("Outputs", "models", "artifacts_three_month_macropru_models.rds"))

rows <- modelsummary_basic(
    macropru_three_month_results$three_lending_models$three_month_implementation,
    coef_omit = 2:16,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Implementation estimates",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

modelsummary_word(
  macropru_three_month_results$three_lending_models$three_month_draft,
  coef_omit = 2:16,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = rows_tbl
) %>%
  width(j = 1, width = 1.5) %>%
  add_body_row(values =
                 c("Draft regulations estimates"),
               colwidths = 10) %>%
  merge_at(i = 3, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporations", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 3) %>% 
  hline(i = 4, j = 1:10, part = "body") 
```

\elandscape


\blandscape

```{r}
#| label: tbl-macropru_draft_rates
#| tbl-cap: "Macroprudential regulation and lending rates"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

rows <- modelsummary_basic(
    macropru_results$rates_models$implementation,
    coef_omit = 2:16,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Implementation estimates",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)


modelsummary_word(
    macropru_results$rates_models$draft,
    coef_omit = 2:16,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
    add_rows = rows_tbl
  ) %>% 
  width(j = 1, width = 1.5) %>% 
   add_body_row(values =
                 c("Draft regulations estimates"),
               colwidths = 10) %>%
  merge_at(i = 3, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporations", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 3) %>% 
  hline(i = 4, j = 1:10, part = "body") 

```

\elandscape


\newpage

## Competition regulation

```{r}
#| label: tbl-competition_unsecured
#| tbl-cap: "Entry and exit effect on unsecured lending"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

competition_results <- read_rds(here("Outputs", "models", "artifacts_competition_models.rds"))


household_rows <- modelsummary_basic(
  competition_results$household_unsecured_model,
  coef_omit = 4:18
) [, c(2, 4:5)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  ) 

household_rows_tbl <- 
  tibble(
  "term" = "Households estimates",  
  "(1)" = " ", 
  "(2)" = " "
  ) %>% 
  add_row(household_rows)



modelsummary_word(
  competition_results$corporate_unsecured_model,
  coef_omit = 4:18,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = household_rows_tbl
) %>%
  add_body_row(values =
                 c("Corporation estimates"),
               colwidths = 3) %>%
  merge_at(i = 5, j = 1:3) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 5) %>%
  hline(i = 8, j = 1:3, part = "body") %>%
  autofit(add_w = 1)

```

\newpage

```{r}
#| label: tbl-competition_corporate_mortgages
#| tbl-cap: "Entry and exit effect on mortgages"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

mortgages_rows <- modelsummary_basic(
  competition_results$households_mortgages_model,
  coef_omit = 4:18
) [, c(2, 4:5)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  )


mortgages_rows_tbl <- 
  tibble(
  "term" = "Households estimates",  
  "(1)" = " ", 
  "(2)" = " "
  ) %>% 
  add_row(mortgages_rows)


modelsummary_word(
  competition_results$commercial_mortgages_model,
  coef_omit = 4:18,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = mortgages_rows_tbl
) %>% 
  add_body_row(values =
                 c("Corporation estimates"),
               colwidths = 3) %>%
  merge_at(i = 5, j = 1:3) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 5) %>%
  hline(i = 8, j = 1:3, part = "body") %>%
  autofit(add_w = 1)


```



\newpage

```{r}
#| label: tbl-competition_corporate_leasing
#| tbl-cap: "Entry and exit effect on leasing and installments"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

leasing_rows <- 
  modelsummary_basic(
  competition_results$household_leasing_installments_model,
  coef_omit = 4:18
) [, c(2, 4:5)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  )

leasing_rows_tbl <- 
  tibble(
  "term" = "Households estimates",  
  "(1)" = " ", 
  "(2)" = " "
  ) %>% 
  add_row(leasing_rows)

modelsummary_word(
  competition_results$coorporate_leasing_installments_model,
  coef_omit = 3:17,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = leasing_rows_tbl
) %>%
  add_body_row(values =
                 c("Corporation estimates"),
               colwidths = 3) %>%
  merge_at(i = 4, j = 1:3) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 4) %>%
  hline(i = 7, j = 1:3, part = "body") %>%
  autofit(add_w = 1)

```


\newpage

```{r}
#| label: tbl-competition_financial_inclusion
#| tbl-cap: "Miscellaneous competition results"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

other_rows <-
  modelsummary_basic(
    competition_results$other_competition_model,
    coef_omit = 3:17,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.") %>% 
  add_row(
    term = "Micro-regulations estimates",
    `(1)` = " ",
    `(2)` = " ",
    `(3)` = " ",
    `(4)` = " ",
    `(5)` = " ",
    `(6)` = " "
  ) %>% 
  add_row(
    modelsummary_basic(
      competition_results$finance_regulations_model,
      coef_omit = 3:17
    )
  ) %>%
  dplyr::select(2, 4:9) %>%
  add_row(
    term = "Bank Fixed Effects",
    `(1)` = "Yes",
    `(2)` = "Yes",
    `(3)` = "Yes",
    `(4)` = "Yes",
    `(5)` = "Yes",
    `(6)` = "Yes"
  ) %>%
  add_row(
    term = "Monthly Fixed Effects",
    `(1)` = "Yes",
    `(2)` = "Yes",
    `(3)` = "Yes",
    `(4)` = "Yes",
    `(5)` = "Yes",
    `(6)` = "Yes"
  )


other_rows_tbl <- 
  tibble(
    "term" = "Other competition developments estimates",
    `(1)` = " ",
    `(2)` = " ",
    `(3)` = " ",
    `(4)` = " ",
    `(5)` = " ",
    `(6)` = " "
  ) %>% 
  add_row(other_rows)

modelsummary_word(
  competition_results$financial_inclusion_model,
  coef_omit = 3:17,
   gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = other_rows_tbl
) %>%
  add_body_row(values =
                 c("Financial inclusion estimates"),
               colwidths = 7) %>%
  merge_at(i = 4, j = 1:7) %>%
  merge_at(i = 7, j = 1:7) %>% 
  align(align = "center", part = "body", i = c(1, 4, 7)) %>%
  hline(i = 9, j = 1:7, part = "body") %>%
  autofit(add_w = 0)

```




# Conclusion

\newpage
# References 
::: {#refs}
:::


\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\newpage

# Appendix

## Data sources

```{r}
#| label: tbl-data
#| tbl-cap: "Data Sources"
#| cache: FALSE
#| include: TRUE
#| ft.arraystrech: .4

tibble(
  " " = c(
    "Macroprudential narrative index",
    "Competition narrative index",
    "BA900",
          "BA930",
          "Controls"),
  "Description" = c(
    " ",
    " ",
    "Banking sector balance sheet data at a bank level",
    "Banking sector lending rates at a bank level",
    "Banking sector performance data at a bank level"
  ),
  
  "Availability" = c(
    "Public data",
    "Public data",
    "Public data",
    "Aggregated data is public.
Bank specific data is private",
"Aggregated data is public.
Bank specific data is private"
  ),
"Source" = c(
  "Own analysis",
  "Own analysis",
  "South African Reserve Bank",
  "South African Reserve Bank",
  "Prudential Authority"
)
) %>%
  flextable_word() %>%
  flextable::align(align = "left", part = "all") %>% 
  autofit(add_w = -0.30)


```


```{r}
#| label: tbl-data1
#| tbl-cap: "Data Description"
#| cache: FALSE
#| include: FALSE
#| ft.arraystretch: 0.5


model_variables <- c( "L_{b,t}",
                      "r_{b,t}",
                      "C - C_{min}",
                      "r_t",
                      "\\Omega_{b,t}"
  
)

tibble(
  "Variable Description" = c("Loans by lending category and bank",
                 "Lending rate by lending category and bank",
                 "Capital buffer",
                 "Capital requirement",
                 "Repo rate",
                 "GDP",
                 "Bank level perfomance metrics"
                 ),
  "Data Description" = c("BA 900 data on bank level credit at a monthly frequency. 
We have summarised to six lending categories: 
Household secured credit, Household unsecured credit, Household residential mortgages, 
Non financial sector secured credit, Non financial sector unsecured credit, 
and Non financial sector mortgages as explain in Section 8.2.",
"BA 930 data on bank level lending rates at a monthly frequency. 
Lending rates are defined as the weighted average rate 
by lending category. These were also summarised into the 
same six lending categories as shown in Section 8.2.",
"Aggregate amount of qualifying capital and 
reserve funds less minimum required capital and reserve funds.",
"Basel III required level of capital as a percentage of risk-weighted assets",
"South African Reserve Bank policy rate",
"Nominal gross domestic product.",
"The following are bank perfomance metrics are include in the data: 
total assets, gross loan advances, retained earnings,
net interest income (12 months), level one high-quality liquid assets required to be held, 
average daily amount of level one high-quality liquid assets held up to 
fourteenth business day of the month following the month to which this return relates, 
aggregate risk weighted exposure, return on equity, return on assets, 
total capital adequacy ratio, and leverage ratio."
                    
                    ),
  "Measure" = c("Rand","Percent", "Percent", "Percent", "Percent", "Rand", "Rand and percent"),
  "Cross-section" = c("Nedbank, First National Bank, Standard Bank, Absa",
                      "Nedbank, First National Bank, Standard Bank, Absa",
                      "Nedbank, First National Bank, Standard Bank, Absa",
                      "Nedbank, First National Bank, Standard Bank, Absa",
                      "NA",
                      "NA",
                      "Nedbank, First National Bank, Standard Bank, Absa"
                      ),
  "Sample" = c("January 2008 to November 2022", 
               "January 2012 to June 2022",
               "January 2008 to September 2020",
               "January 2008 to September 2020",
               "January 2008 to February 2021",
               "March 2008 to March 2022",
               "January 2008 to September 2022"),
  "Frequency" = c (
  "Monthly",
  "Monthly",
  "Monthly",
   "Monthly",
  "Monthly",
  "Quartely",
  "Monthly"
  ),
  
) %>% 
  flextable_word(
    caption = "Data Description") %>%
  align(j = 1, part = "all", align = "left") %>% 
  fit_to_width(max_width = 10) 

```

\newpage
## Aggregation scheme

```{r}
#| label: tbl-scheme
#| tbl-cap: "Aggregation schema"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: .55

tibble(
  "BA 900 Categories" = 
    c(
      "Installment sales",
      "",
      "",
      "",
      "Leasing transactions",
      "",
      "",
      "",
      "Farm mortgages",
      "",
      "",
      "Residential mortgages",
      "",
      "",
      "Commercial and other mortgages",
      "",
      "",
      "",
      "",
      "",
      "Credit cards",
      "",
      "",
      "",
      "Overdrafts",
      "",
      "",
      "",
      "",
      "",
      "Factoring debtors",
      "Other loans and advances", 
      "",
      "", 
      "",
      "",
      ""
    ),
  "Item Number" = c(
    141,
    142,
    143,
    144,
    146,
    147,
    148,
    149,
    152,
    153,
    154,
    156,
    157,
    158,
    160,
    161,
    162,
    163,
    164,
    165,
    167,
    168,
    169,
    170,
    178,
    181,
    182,
    183,
    184,
    185,
    187,
    189,
    190, 
    191, 
    192,
    193,
    ""
    
  ),
  "Sector" = c(
    "Financial corporate sector",
    "Non financial corporate sector",
    "Household sector",
    "Other",
    "Financial corporate sector",
    "Non financial corporate sector",
    "Household sector",
    "Other",
    "Non financial corporate sector",
    "Household sector",
    "Other",
    "Non financial corporate sector",
    "Household sector",
    "Other",
    "Public financial corporates",
    "Public non-financial corporates",
    "Private financial corporate",
    "Private non-financial corporates",
    "Household sector",
    "Other",
    "Financial corporate sector",
    "Non financial corporate sector",
    "Household sector",
    "Other",
    "Public sector (includes public corporations and local government)",
    "Financial corporate sector",
    "Non financial corporate sector",
    "Unincorporated business enterprises",
    "Other Household sector",
    "Non-profit organisations serving households",
    "",
    "Financial corporate sector",
    "Non financial corporate sector",
    "Unincorporated business enterprises",
    "Other Household sector",
    "Non-profit organisations serving households",
    ""
  ),
  "Aggregation Key" = c(
    "-",
    "g",
    "h",
    "-",
    "-",
    "g",
    "h",
    "-",
    "d",
    "d",
    "-",
    "e",
    "-",
    "-",
    "-",
    "-",
    "d",
    "d",
    "-",
    "-",
    "a",
    "b",
    "-",
    "-",
    "-",
    "-",
    "-",
    "a",
    "-",
    "b",
    "-",
    "-",
    "a",
    "-",
    "b",
    "-",
    ""
  ),
) %>% 
  flextable_word(caption = "Aggregation schema") %>% 
  align(align = "left", part = "all") %>%
  colformat_num(digits = 0, j = 2) %>% 
  autofit(add_w = -0.2)


```


The following aggregation scheme which results in nine categories was followed based on @tbl-scheme:

a.	Non-financial corporate unsecured lending: Items 168 + 183 + 190 
b.	Household unsecured lending: Items 169 + 185 + 192   
c.	Total unsecured lending: Non-financial corporate unsecured lending + Household unsecured lending
d.	Commercial mortgages to corporates and households: Items 152 + 153 + 156 + 163 + 164   
e.  Residential mortgages to household: Item 157
f.	Total mortgage lending: Commercial mortgages to corporates and households + Residential mortgages to household
g.	Leasing and instalments to corporates: Items 142 + 147
h.  Leasing and instalments to households: Items 143 + 148
i.  Total leasing and instalments: Leasing and instalments to corporates + Leasing and instalments to households

\newpage
## Bank lending rates weighting scheme

The loans quantities from the BA900s are then linked to the lending rate data from the BA930s using table to create nine lending rate categories the schema on @tbl-scheme_1. The weights for each category are then calculated by dividing the total value of the loans in each category by the total value of all loans in the BA900s. The weights are then used to calculate the weighted average lending rate for each month. The weighted average lending rate is calculated by multiplying the lending rate for each category by the weight for that category and then summing the results.

```{r}
#| label: tbl-scheme_1
#| tbl-cap: "Weighting schema"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: .55

tibble(
  "Sector" = c(
   "Corporate sector",
   "",
   "",
   "",
   "",
   "",
   "",
   "",
   "",
   "Household sector",
   "",
   "",
   "",
   "",
   "",
   "",
   "",
   ""
  ),
  "BA 930 Categories" = c(
    "Overdraft rate",
    "Instalment sale agreements flexible rate",
    "Instalment sale fixed rate",
    "Leasing transactions flexible rate",
    "Leasing transactions fixed rate",
    "Mortgage advances flexible rate",
    "Mortgage advances fixed rate",
    "Credit card rate",
    "Other",
    "Overdraft rate",
    "Instalment sale agreements flexible rate",
    "Instalment sale fixed rate",
    "Leasing transactions flexible rate",
    "Leasing transactions fixed rate",
    "Mortgage advances flexible rate",
    "Mortgage advances fixed rate",
    "Credit card rate",
    "Other"
  ),
  "Item Number" = c(
    48,
    49,
    50,
    51,
    52,
    53,
    54,
    55,
    56,
    58,
    59,
    60,
    61,
    62,
    63,
    64,
    65,
    66
  ),
  "Weighting Key" = c(
    "a and c",
    "g and i",
    "-",
    "g and i",
    "-",
    "d and f",
    "-",
    "a and c",
    "a and c",
    "b and c",
    "h and i",
    "-",
    "h and i",
    "-",
    "e and f",
    "-",
    "b and c",
    "b and c"
  )
) %>% 
  flextable_word() %>% 
  align(align = "left", part = "all") %>%
  colformat_num(digits = 0, j = 3) %>% 
  autofit(add_w = .35)

```

The nine categories, therefore, are as follows:

a.	Non-financial corporate unsecured lending: Weighted average of items 55 + 48 + 56
b.	Household unsecured lending: Weighted average of items 65 + 58 + 66
c.	Total unsecured lending: Weighted average of items 55 + 48 + 56 + 65 + 58 + 66
d.	Commercial mortgages to corporates and households: Weighted average of items 53
e.  Residential mortgages to household:  Item 63
f.	Total mortgage lending: Weighted average of items 53 + 63
g.	Leasing and instalments to corporates: Weighted average of items 49 + 51
h.  Leasing and instalments to households: Weighted average of items 59 + 61
i.  Total leasing and instalments: Weighted average of items 49 + 51 + 59 + 61




\newpage
## Aggregated bank lending

```{r}
#| label: fig-bank_lending
#| fig-cap: "Total aggregated bank lending"
#| fig-height: 6
#| fig-width: 6
#| cache: TRUE

bank_lending <- read_rds(here("Outputs","BA900","artifacts_BA900.rds"))
bank_lending_gg <- bank_lending$aggregated_graphs$total_aggregation_gg
bank_lending_gg
```


## Weighted lending rates (aggregated)

```{r}
#| label: fig-bank_interest_rates
#| fig-cap: "Weighted lending rates"
#| fig-height: 5
#| fig-width: 6
#| cache: FALSE

bank_interest_rates <- 
  read_rds(here("Outputs","BA930","artifacts_BA930_futher_analysis.rds"))

bank_interest_rates_gg <- bank_interest_rates$gg$aggregated_lending_rates_gg
bank_interest_rates_gg
```


<!-- ## Aggregated bank controls (check) -->

```{r}
#| label: fig-controls
#| fig-cap: "Aggregated bank controls"
#| fig-height: 7.5
#| fig-width: 6
#| cache: FALSE
#| include: FALSE
#| eval: FALSE

controls <- read_rds(here("Outputs","BA920","artifacts_BA920.rds"))
controls_gg <- controls$graphs$Total_banks_gg
controls_gg
```




\newpage
## Description of narrative events

```{r}
#| label: tbl-description_narratives
#| tbl-cap: "Description of the narrative indexes"
#| cache: FALSE
#| include: TRUE

tibble(
  "Index" = c(
    "Macroprudential",
    "Competition"
  ),
  "Description" = c(
    "The macroprudential narrative index",
    "The competition narrative index"
  )
) %>% 
  flextable_word() %>% 
  align(align = "left", part = "all") %>%
  autofit(add_w = 1.5)
```


\newpage
## Macroprudential narrative indexes

```{r}
#| label: fig-macro_narrative_indexes
#| fig-cap: "Macroprudential narrative indexes"
#| fig-height: 3
#| fig-width: 6
#| cache: FALSE

macro_narratives <- read_rds(here("Outputs","narratives","artifacts_macropru_narratives.rds"))
macropru_narratives_gg <- macro_narratives$macropru_narratives_gg
macropru_narratives_gg

```

\newpage

## Competition narrative indexes

```{r}
#| label: fig-comp_narrative_indexes
#| fig-cap: "Competition narrative indexes"
#| fig-height: 7
#| fig-width: 6
#| cache: FALSE

comp_narratives <- read_rds(here("Outputs","narratives","artifacts_competion_narratives.rds"))

comp_narratives_gg <- comp_narratives$graphs$combined_dummies_gg
comp_narratives_gg
```












## Macropudential results with controls

\blandscape

```{r}
#| label: tbl-macropru_draft_lending_control
#| tbl-cap: "Macroprudential regulations and lending volumes (3-months) with controls results"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

macropru_control_results <- read_rds(here("Outputs", "models", "artifacts_macropru_controls_models.rds"))

rows <- modelsummary_basic(
    macropru_control_results$lending_models$implementation,
    coef_omit = 4:17,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Implementation estimates",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

modelsummary_word(
  macropru_control_results$lending_models$draft,
  coef_omit = 4:17,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = rows_tbl
) %>%
  width(j = 1, width = 1.8) %>%
  add_body_row(values =
                 c("Draft regulations estimates"),
               colwidths = 10) %>%
  merge_at(i = 5, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporations", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 5) %>% 
  hline(i = 8, j = 1:10, part = "body") 
```

\elandscape


\blandscape

```{r}
#| label: tbl-macropru_draft_rates_control
#| tbl-cap: "Macroprudential regulation and lending rates with controls"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

rows <- modelsummary_basic(
    macropru_control_results$rates_models$implementation,
    coef_omit = 4:17,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Implementation estimates",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)


modelsummary_word(
    macropru_control_results$rates_models$draft,
    coef_omit = 4:17,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
    add_rows = rows_tbl
  ) %>% 
  width(j = 1, width = 1.8) %>% 
   add_body_row(values =
                 c("Draft regulations estimates"),
               colwidths = 10) %>%
  merge_at(i = 5, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporations", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 5) %>% 
  hline(i = 8, j = 1:10, part = "body") 

```

\elandscape

## Competition results with controls

```{r}
#| label: tbl-competition_unsecured_controls
#| tbl-cap: "Entry and exit effect on unsecured lending with controls"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

competition_control_results <- read_rds(here("Outputs", "models", "artifacts_competition_control_models.rds"))


household_rows <- modelsummary_basic(
  competition_control_results$household_unsecured_control_model,
  coef_omit = 8:21
) [, c(2, 4:5)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  ) 

household_rows_tbl <- 
  tibble(
  "term" = "Households estimates",  
  "(1)" = " ", 
  "(2)" = " "
  ) %>% 
  add_row(household_rows)

modelsummary_word(
  competition_control_results$corporate_unsecured_control_model,
  coef_omit = 8:21,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = household_rows_tbl
) %>%
  add_body_row(values =
                 c("Corporation estimates"),
               colwidths = 3) %>%
  merge_at(i = 9, j = 1:3) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 9) %>%
  hline(i = 16, j = 1:3, part = "body") %>%
  autofit(add_w = 1)

```

\newpage

```{r}
#| label: tbl-competition_corporate_mortgages_controls
#| tbl-cap: "Entry and exit effect on mortgages with controls"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

mortgages_rows <- modelsummary_basic(
  competition_control_results$households_mortgages_control_model,
  coef_omit = 8:21,
) [, c(2, 4:5)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  )


mortgages_rows_tbl <- 
  tibble(
  "term" = "Households estimates",  
  "(1)" = " ", 
  "(2)" = " "
  ) %>% 
  add_row(mortgages_rows)


modelsummary_word(
  competition_control_results$commercial_mortgages_control_model,
  coef_omit = 8:21,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = mortgages_rows_tbl
) %>% 
  add_body_row(values =
                 c("Corporation estimates"),
               colwidths = 3) %>%
  merge_at(i = 9, j = 1:3) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 9) %>%
  hline(i = 16, j = 1:3, part = "body") %>%
  autofit(add_w = 1)


```


\newpage

```{r}
#| label: tbl-competition_corporate_leasing_controls
#| tbl-cap: "Entry and exit effect on leasing and installments with controls"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

leasing_rows <- 
  modelsummary_basic(
  competition_control_results$household_leasing_installments_control_model,
  coef_omit = 8:21,
) [, c(2, 4:5)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes"
  )

leasing_rows_tbl <- 
  tibble(
  "term" = "Households estimates",  
  "(1)" = " ", 
  "(2)" = " "
  ) %>% 
  add_row(leasing_rows)

modelsummary_word(
  competition_control_results$coorporate_leasing_installments_control_model,
  coef_omit = 7:20,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = leasing_rows_tbl
) %>%
  add_body_row(values =
                 c("Corporation estimates"),
               colwidths = 3) %>%
  merge_at(i = 8, j = 1:3) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 8) %>%
  hline(i = 15, j = 1:3, part = "body") %>%
  autofit(add_w = 1)

```


\newpage

```{r}
#| label: tbl-competition_financial_inclusion_controls
#| tbl-cap: "Miscellaneous competition results with controls"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

other_rows <-
  modelsummary_basic(
    competition_control_results$other_competition_control_model,
    coef_omit = 7:20,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.") %>% 
  add_row(
    term = "Micro-regulations estimates",
    `(1)` = " ",
    `(2)` = " ",
    `(3)` = " ",
    `(4)` = " ",
    `(5)` = " ",
    `(6)` = " "
  ) %>% 
  add_row(
    modelsummary_basic(
      competition_control_results$finance_regulations_control_model,
      coef_omit = 7:20
    )
  ) %>%
  dplyr::select(2, 4:9) %>%
  add_row(
    term = "Bank Fixed Effects",
    `(1)` = "Yes",
    `(2)` = "Yes",
    `(3)` = "Yes",
    `(4)` = "Yes",
    `(5)` = "Yes",
    `(6)` = "Yes"
  ) %>%
  add_row(
    term = "Monthly Fixed Effects",
    `(1)` = "Yes",
    `(2)` = "Yes",
    `(3)` = "Yes",
    `(4)` = "Yes",
    `(5)` = "Yes",
    `(6)` = "Yes"
  )


other_rows_tbl <- 
  tibble(
    "term" = "Other competition developments estimates",
    `(1)` = " ",
    `(2)` = " ",
    `(3)` = " ",
    `(4)` = " ",
    `(5)` = " ",
    `(6)` = " "
  ) %>% 
  add_row(other_rows)

modelsummary_word(
  competition_control_results$financial_inclusion_control_model,
  coef_omit = 7:20,
   gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = other_rows_tbl
) %>%
  add_body_row(values =
                 c("Financial inclusion estimates"),
               colwidths = 7) %>%
  merge_at(i = 8, j = 1:7) %>%
  merge_at(i = 15, j = 1:7) %>% 
  align(align = "center", part = "body", i = c(1, 8, 15)) %>%
  hline(i = 21, j = 1:7, part = "body") %>%
  autofit(add_w = 0)

```




