---
title: "Regulation and bank lending in South Africa: A narrative approach"
author: "Dumakude Nxumalo and Keaoleboga Mncube"
institute: "University of Pretoria"
date: "28 March 2024"
format: 
    revealjs:
      theme: simple
      show-slide-number: all
      controls: true
      progress: true
      center: false
      # width: 1
      # height: 1080
      # aspectratio: 1
      number-sections: false
      center-title-slide: true
      title-slide-style: default
      auto-strech: true
      margin: 0.2
      reference-location: document
---
```{r setup}
#| include=FALSE
options(scipen=999)
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      dpi = 500, 
                      fig.pos = "H", 
                      fig.env= 'figure', 
                      tab.topcaption = TRUE,
                      ft.align = "center",
                      ft.arraystretch = 1
                      )
library(tidyverse)
library(readr)
library(readxl)
library(zoo)
library(magick)
library(stringi)
library(patchwork)
library(xtable)
library(scales)
library(rvest)
library(rebus)
library(PNWColors)
library(lubridate)
library(here)
library(flextable)
library(gt)
library(officer)
library(officedown)
library(modelsummary)
library(sandwich)
library(lmtest)

set_flextable_defaults(
  font.size = 10,
  digits = 3,
  font.family = "Arial",
  border.color = "#5d74a5",
  border.width = 1,
  fp_border = fp_border(color = "#5d74a5", width = 1, style = "solid")
)


```

```{r functions}
source(here("Functions","flextable_word.R" ))
source(here("Functions","modelsummary_word.R" ))
source(here("Functions","modelplot_word.R" ))
source(here("Functions","borderlines.R" ))
```

## Introduction
- Since the global financial crisis, regulatory changes have been a key focus for policymakers^[National Treasury. 2018. Explanatory Policy Paper for COFI Bill.].
- However, there are concerns as to the impact of these changes on bank lending. This has been a focus of various studies.^[Aiyar et al. 2016. How does credit supply respond to monetary and bank minimum capital requirements? European Economic Review. Other studies as well.]
- Given the importance of bank lending to economic growth, understanding the factors that influence bank lending is crucial.^[Loayza et al. 2017. Financial development, growth, and crisis: Is there a trade off? World Bank.]

## The aim of the study
- Therefore, this study aims to investigate the impact of regulatory changes on bank lending in South Africa using a narrative approach.
- We focus on two types of regulatory changes:
  - Macroprudential regulation
  - Finance and inclusion regulation
- Estimate panel regression for six credit categories and four banks

## Brief comments on literature
- The literature on the impact of regulatory changes on bank lending is diverse.
- Some studies find that regulatory changes have a negative impact on bank lending. Whilst other suggest that regulation does not impact bank lending.
- In South Africa studies are limited, and therefore this question remains open.

# Methodology

## Narrative approach  {.smaller}

- The narrative approach requires construction of 'exogenous' narrative indices. Exogeneity is a key challenge in the literature.
- The indices reflect regulatory events which the literature identifies as having a potential impact on bank lending.
- However, the narrative approach is not without its limitations.
- For example, the indices do not conform to any a priori expectations which can make interpretation challenging.
- However, the narrative approach is useful in that it allows for the identification of regulatory events that may have been overlooked in the literature.

## Construction of macroprudential indices
- We delineate between anticipation and actual effect of the regulation
- The anticipation effects is captured by the communication of macroprudential regulation,
- Therefore, $Draft_t$ take a value of 1 when there is communication and 0 otherwise.
- Similarly, the actual impact if captured using the $Implementation_t$ variable which take a value of 1 when macroprudential regulation is implemented, and 0 otherwise

## Construction of finance indices
- Following the same logic as the macroprudential indices, we construct two finance indices. 
- The first index captures proposed and final changes in finance regulation ($FinReg_t$), which take a value of 1 when such an event occurs, and 0 otherwise.
- The second index captures the proposed national financial inclusion framework ($FinInclusion_t$), which takes a value of 1 when this event occurs, and 0 otherwise.

## Construction of credit categories (segments)
- Using the BA900, 930, and other data sources we constructed six credit categories.
- Households: mortgages, unsecured credit, and secured credit
- Corporate: mortgages, unsecured credit, and secured credit
- Aggregates: total mortgages, total unsecured credit, total secured credit.

## Econometric approach
- We use a panel data approach to estimate the impact of regulatory changes on bank lending.
- For each index the fixed effect model specified as follows:
$$ Lending_{b,t}^{c} = \beta_1Index_t^i + \beta_2\tau_b^c + \beta_3\lambda_t^c + \beta_4\Omega^c + \epsilon_{b,t}^c, and  $$
$$ Rate_{b,t}^{c} = \alpha_1Index_t^i + \alpha_2\tau_b^c + \alpha_3\lambda_t^c + \alpha_4\Omega^c + \epsilon_{b,t}^c .$$
- Where $b$ is FNB, Absa, Nedbank, or Standard Bank. $t$ is the time period, $c$ is the credit category, and $i$ constitutes the $Draft_t$, $Implementation_t$, $FinReg_t$, or $FinInclusion_t$ indices. 

## Econometric approach (continued)
- Therefore $Lending_{b,t}^c$ are the bank level three month growth of lending, and $Rate_{b,t}^c$ are the bank level lending rates.
- $\Omega^c$ is a matrix of controls which includes a covid dummy, bank level variables, and other macroeconomic variables.
- $\tau_b^c$ captures the bank fixed effect, $\lambda_t^c$ are the time fixed effects, and $\epsilon_{b,t}^c$ are the error terms.

# Data
## Bank lending
```{r}
#| label: fig-bank_lending
#| fig-cap: "Total aggregated bank lending"
#| fig-height: 6
#| fig-width: 12
#| cache: TRUE

bank_lending <- read_rds(here("Outputs","BA900","artifacts_BA900.rds"))
bank_lending_gg <- bank_lending$aggregated_graphs$total_aggregation_gg
bank_lending_gg +
  facet_wrap (. ~ Series, scale = "free", ncol = 3) +
  theme(text = element_text(size = 9, face = "bold"))
```

## Bank lending rates
```{r}
#| label: fig-bank_interest_rates
#| fig-cap: "Weighted lending rates"
#| fig-height: 6
#| fig-width: 10
#| cache: FALSE

bank_interest_rates <- 
  read_rds(here("Outputs","BA930","artifacts_BA930_futher_analysis.rds"))

bank_interest_rates_gg <- bank_interest_rates$gg$aggregated_lending_rates_gg
bank_interest_rates_gg +
  facet_wrap (. ~ Series, scale = "free", ncol = 3) +
   theme(text = element_text(size = 9, face = "bold")) 
  
```

## Macroprudential regulation narrative indices
```{r}
#| label: fig-macro_narrative_indexes_three_month
#| fig-cap: "Three month lending growth and macroprudential narrative indexes comparison. Note: The black line represents the Draft index, and the red line represents the Implementation index."
#| fig-height: 8
#| fig-width: 20
#| cache: FALSE

narratives <- read_rds(here("Outputs","combined_data", "artifacts_visual_lag_test.rds"))
macropru_narratives_one_month_gg <- narratives$macropru$three_month_test_macropru_gg
macropru_narratives_one_month_gg +
  facet_wrap(~series, scales = "free_y", ncol = 3) +
  theme(text = element_text(size = 14, face = "bold")
        )

```

## Macroprudential regulation narrative indices (examples)
- For example:
  - 2013 implementation of BASEL III
  - Guidance note II of 2012 - new definition of regulatory capital
  - Guidance note 6 of 2013 - only equities listed on JSE can be included in level 1 capital

## Finance regulation narrative indices
```{r}
#| label: fig-comp_narrative_indexes_three_month
#| fig-cap: "Three month lending growth and financial narrative indexes comparison. Note: The black line represents the Financial regulation index, and the red line represents the Financial inclusion index."
#| fig-height: 8
#| fig-width: 20
#| cache: FALSE

finance_narratives_three_months_gg <- narratives$finance$three_month_test_finance_gg
finance_narratives_three_months_gg +
   facet_wrap(~series, scales = "free_y", ncol = 3) +
  theme(text = element_text(size = 14, face = "bold")
        )
```
## Finance regulation narrative indices (examples)
- For example:
  - Removal of adverse customer credit information (2013/14)
  - Reduction in the max interest on credit that could offered by lenders (2015)
  - National credit regulations amendments (2014-15)

# Macroprudential results

## Macropru regulation and lending volumes
```{r}
#| label: tbl-macropru_draft_lending
#| tbl-cap: "Macroprudential regulations and lending volumes (3-months) results"
#| cache: FALSE
#| include: TRUE
#| eval: TRUE
#| ft.arraystretch: 1


macropru_results <- read_rds(here("Outputs", "models", "artifacts_macropru_models.rds"))

macropru_three_month_results <- read_rds(here("Outputs", "models", "artifacts_three_month_macropru_models.rds"))

rows <- modelsummary_basic(
    macropru_three_month_results$three_lending_models$three_month_implementation,
    coef_omit = 2:16,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Implementation model",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

predictor_macro <- c("Draft", "Implementation")

dependant_lending <- c(
  " ",
  "Unsecured",
  "Secured",
  "Mortgages",
  "Unsecured",
  "Secured",
  "Mortgages",
  "Unsecured",
  "Secured",
  "Mortgages"
)
modelsummary_word(
  macropru_three_month_results$three_lending_models$three_month_draft,
  coef_omit = 2:16,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = rows_tbl
) %>%
  width(j = 1, width = 3) %>%
  add_body_row(values =
                 c("Draft model"),
               colwidths = 10) %>%
  merge_at(i = 3, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporates", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 3) %>%
  hline(i = 4, j = 1:10, part = "body") %>%
  mk_par(
    i = 2,
    part = "header",
    value = as_paragraph(dependant_lending)
  ) %>% 
  mk_par(
    j = 1,
    i = c(2, 4),
    part = "body",
    value = as_paragraph(predictor_macro)
  ) %>% 
  hline(i = 1, j = 1:1, part = "header", border = fp_border(color = "white"))
```

## Macropru regulation and lending rates 
```{r}
#| label: tbl-macropru_draft_rates
#| tbl-cap: "Macroprudential regulation and lending rates"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

rows <- modelsummary_basic(
    macropru_results$rates_models$implementation,
    coef_omit = 2:16,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Implementation model",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

dependant_rate <- c(
  " ",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage"
)



modelsummary_word(
    macropru_results$rates_models$draft,
    coef_omit = 2:16,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
    add_rows = rows_tbl
  ) %>% 
  width(j = 1, width = 3) %>% 
   add_body_row(values =
                 c("Draft model"),
               colwidths = 10) %>%
  merge_at(i = 3, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporates", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 3) %>% 
  hline(i = 4, j = 1:10, part = "body") %>% 
   mk_par(
    i = 2,
    part = "header",
    value = as_paragraph(dependant_rate)
  ) %>% 
  mk_par(
    j = 1,
    i = c(2, 4),
    part = "body",
    value = as_paragraph(predictor_macro)
  ) %>% 
  hline(i = 1, j = 1:1, part = "header", border = fp_border(color = "white")) 

```

## Discussion - Macroprudential regulation
- The macroprudential regulation results suggests lending portfolio rebalancing phenomena. This result was also robust with controls.
- In terms of the lending volumes results:
  - There is a rebalancing from long to short term lending. 
  - Suggesting that banks are more risk averse toward term lending due to regulatory requirements
  
## Discussion - Macroprudential regulation (continued)
- In terms of the lending rates results:
  - Rebalancing means banks shift from unsecured credit to secured credit. This effect was not prevalent in mortgages.
  - The rebalancing effect was more pronounced in the household sector than in the corporate sector.
  - This is a similar risk averse behavior as in the lending volumes results.

# Finance and inclusion regulation results
## Finance regulation and lending volumes
```{r}
#| label: tbl-finance_lending
#| tbl-cap: "Finance regulation and lending volumes (3-months) results"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

additional_competition_results <- read_rds(here("Outputs", "models", "artifacts_additional_competition_models.rds"))

rows <- modelsummary_basic(
    additional_competition_results$financial_inclusion_lending_models,
    coef_omit = 3:17,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Financial inclusion model",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

dependant_rate <- c(
  " ",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage"
)

predictor_finance <- c("FinReg", "Covid-dummy", "FinInclusion", "Covid-dummy")

modelsummary_word(
    additional_competition_results$finance_lending_models,
    coef_omit = 3:17,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
    add_rows = rows_tbl
  ) %>% 
  width(j = 1, width = 3) %>% 
   add_body_row(values =
                 c("Finance regulation model"),
               colwidths = 10) %>%
  merge_at(i = 4, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporates", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 4) %>% 
  hline(i = 6, j = 1:10, part = "body") %>% 
   mk_par(
    i = 2,
    part = "header",
    value = as_paragraph(dependant_rate)
  ) %>% 
  mk_par(
    j = 1,
    i = c(2, 3, 5, 6),
    part = "body",
    value = as_paragraph(predictor_finance)
  ) %>% 
  hline(i = 1, j = 1:1, part = "header", border = fp_border(color = "white")) 

```

## Finance regulation and lending rates
```{r}
#| label: tbl-finance_rates
#| tbl-cap: "Financial regulation and lending rates results"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

rows <- modelsummary_basic(
    additional_competition_results$financial_inclusion_rates_models,
    coef_omit = 3:17,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Financial inclusion model",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

dependant_rate <- c(
  " ",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage"
)

modelsummary_word(
    additional_competition_results$finance_rates_models,
    coef_omit = 3:17,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
    add_rows = rows_tbl
  ) %>% 
  width(j = 1, width = 3) %>% 
   add_body_row(values =
                 c("Finance regulation model"),
               colwidths = 10) %>%
  merge_at(i = 4, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporates", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 4) %>% 
  hline(i = 6, j = 1:10, part = "body") %>% 
   mk_par(
    i = 2,
    part = "header",
    value = as_paragraph(dependant_rate)
  ) %>% 
  mk_par(
    j = 1,
    i = c(2, 3, 5, 6),
    part = "body",
    value = as_paragraph(predictor_finance)
  ) %>% 
  hline(i = 1, j = 1:1, part = "header", border = fp_border(color = "white"))

```



## Discussion - Finance and inclusion regulation  {.smaller}
- The finance and inclusion regulation results indicate a decrease in lending volumes and an increase in lending rates. However, the lending rates results are sensitive to controls.
- Finance regulation:
  - The lending volumes specifically decrease in the unsecured lending for both corporates and households. 
  - Mortgage lending volumes increase for households
  - For households and corporates lending rates increased for the secured credit, and an increase in lending rates for households mortgages.


## Discussion - Finance and inclusion regulation (continued)  

- Finance inclusion:
  - Similar to finance regulation results but with stronger effects on lending volumes.
- Overall and contrary to the intentions of the changes of these regulations, the results indicate a decrease in lending volumes and an increase in lending rates.


## Conclusions


# Appendix: Results with controls

## Macropru regulation and lending volume
```{r}
#| label: tbl-macropru_draft_lending_control
#| tbl-cap: "Macroprudential regulations and lending volumes (3-months) with controls results"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

macropru_control_results <- read_rds(here("Outputs", "models", "artifacts_macropru_controls_models.rds"))

rows <- modelsummary_basic(
    macropru_control_results$lending_models$implementation,
    coef_omit = 4:18,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Implementation model",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

predictor_control <-
  c(
    "Draft",
    "Return on assets",
    "Total capital adequacy ratio",
    "Implementation",
    "Return on assets",
    "Total capital adequacy ratio"
  )

predictor_control <- c(
  "Draft",
  "Return on assets",
  "Total capital adequacy ratio",
  "Implementation",
  "Return on assets",
  "Total capital adequacy ratio"
)

modelsummary_word(
  macropru_control_results$lending_models$draft,
  coef_omit = 4:18,
  gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
  add_rows = rows_tbl
) %>%
  width(j = 1, width = 3) %>%
  add_body_row(values =
                 c("Draft model"),
               colwidths = 10) %>%
  merge_at(i = 5, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporates", "Households"),
    colwidths = c(1, 3, 3, 3)) %>%
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 5) %>%
  hline(i = 8, j = 1:10, part = "body") %>%
  mk_par(i = 2,
         part = "header",
         value = as_paragraph(dependant_lending)) %>%
  mk_par(
    j = 1,
    i = c(2, 3, 4, 6, 7, 8),
    part = "body",
    value = as_paragraph(predictor_control)
  ) %>% 
  hline(i = 1, j = 1:1, part = "header", border = fp_border(color = "white")) 
```
## Macropru regulation and lending rates 
```{r}
#| label: tbl-macropru_draft_rates_control
#| tbl-cap: "Macroprudential regulation and lending rates with controls results"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

rows <- modelsummary_basic(
    macropru_control_results$rates_models$implementation,
    coef_omit = 4:18,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Implementation model",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)


modelsummary_word(
    macropru_control_results$rates_models$draft,
    coef_omit = 4:18,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
    add_rows = rows_tbl
  ) %>% 
  width(j = 1, width = 3) %>% 
   add_body_row(values =
                 c("Draft model"),
               colwidths = 10) %>%
  merge_at(i = 5, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporations", "Households"),
    colwidths = c(1, 3, 3, 3)) %>%
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 5) %>%
  hline(i = 8, j = 1:10, part = "body") %>%
  mk_par(i = 2,
         part = "header",
         value = as_paragraph(dependant_rate)) %>%
  mk_par(
    j = 1,
    i = c(2, 3, 4, 6, 7, 8),
    part = "body",
    value = as_paragraph(predictor_control)
  ) %>% 
  hline(i = 1, j = 1:1, part = "header", border = fp_border(color = "white"))

```

## Finance regulation and lending volume
```{r}
#| label: tbl-finance_control_lending
#| tbl-cap: "Finance regulation and lending volumes (3-months) with controls results"
#| cache: FALSE
#| include: TRUE

additional_competition_controls_results <- read_rds(here("Outputs", "models", "artifacts_additional_competition_control_models.rds"))

rows <- modelsummary_basic(
    additional_competition_controls_results$financial_inclusion_lending_models,
    coef_omit = 5:19,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Financial inclusion model",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

dependant_rate <- c(
  " ",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage"
)

predictor_finance_controls <-
  c("FinReg",
    "Covid-dummy",
    "Repo rate",
    "Consumer confidence",
    "Return on assets",
    "SAVIT40 index",
    "FinInclusion",
    "Covid-dummy",
    "Repo rate",
    "Consumer confidence",
    "Return on assets",
    "SAVIT40 index")


modelsummary_word(
    additional_competition_controls_results$finance_lending_models,
    coef_omit = 5:19,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
    add_rows = rows_tbl
  ) %>% 
  width(j = 1, width = 5) %>% 
   add_body_row(values =
                 c("Finance regulation model"),
               colwidths = 10) %>%
  merge_at(i = 8, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporates", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 8) %>% 
  hline(i = 14, j = 1:10, part = "body") %>% 
   mk_par(
    i = 2,
    part = "header",
    value = as_paragraph(dependant_rate)
  ) %>% 
  mk_par(
    j = 1,
    i = c(2, 3, 4, 5, 6, 7, 9, 10, 11, 12, 13, 14),
    part = "body",
    value = as_paragraph(predictor_finance_controls)
  ) %>% 
  hline(i = 1, j = 1:1, part = "header", border = fp_border(color = "white")) %>% 
  height_all(height = 0.01, part = "all") %>% 
  hrule(rule = "exact")

```
## Finance regulation and lending rates
```{r}
#| label: tbl-finance_control_rates
#| tbl-cap: "Finance regulation and lending rates with controls results"
#| cache: FALSE
#| include: TRUE
#| ft.arraystretch: 1

rows <- modelsummary_basic(
    additional_competition_controls_results$financial_inclusion_rates_models,
    coef_omit = 5:19,
    gof_omit = "AIC|BIC|Std|Log",
    output = "dataframe"
  ) [, c(2, 4:12)] %>% 
  add_row(
   term = "Bank Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) %>% 
  add_row(
   term = "Monthly Fixed Effects",  
  `(1)` = "Yes", 
  `(2)` = "Yes", 
  `(3)` = "Yes", 
  `(4)` = "Yes", 
  `(5)` = "Yes", 
  `(6)` = "Yes", 
  `(7)` = "Yes", 
  `(8)` = "Yes", 
  `(9)` = "Yes"
  ) 
  
rows_tbl <-  tibble(
  "term" = "Financial inclusion model",  
  "(1)" = " ", 
  "(2)" = " ", 
  "(3)" = " ", 
  "(4)" = " ", 
  "(5)" = " ", 
  "(6)" = " ", 
  "(7)" = " ", 
  "(8)" = " ", 
  "(9)" = " ") %>% 
  add_row(rows)

dependant_rate <- c(
  " ",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage",
  "Unsecured",
  "Secured",
  "Mortgage"
)

modelsummary_word(
    additional_competition_controls_results$finance_rates_models,
    coef_omit = 5:19,
    gof_omit = "AIC|BIC|Std|Log|Num.Obs.",
    add_rows = rows_tbl
  ) %>% 
  width(j = 1, width = 3) %>% 
   add_body_row(values =
                 c("Finance regulation model"),
               colwidths = 10) %>%
  merge_at(i = 8, j = 1:10) %>%
  add_header_row(
    values = c("", "Total", "Corporates", "Households"),
    colwidths = c(1, 3, 3, 3)) %>% 
  align(align = "center", part = "header", i = 1) %>%
  align(align = "center", part = "body", i = 1) %>%
  align(align = "center", part = "body", i = 8) %>% 
  hline(i = 14, j = 1:10, part = "body") %>% 
   mk_par(
    i = 2,
    part = "header",
    value = as_paragraph(dependant_rate)
  ) %>% 
  mk_par(
    j = 1,
    i = c(2, 3, 4, 5, 6, 7, 9, 10, 11, 12, 13, 14),
    part = "body",
    value = as_paragraph(predictor_finance_controls)
  ) %>% 
  hline(i = 1, j = 1:1, part = "header", border = fp_border(color = "white"))

```
